### 2025-04-03 code/eda-code/eda_manuscript.qmd

---
title: " EDA_ Manuscript"
date: "2025-02-23"
output: html_document
---

```{r}
# Load Required Libraries
# Essential libraries for data handling, visualization, and analysis

library(here)             # Manage file paths
library(dplyr)            # Data manipulation
library(tidyverse)        # Data wrangling and visualization
library(janitor)          # Clean column names
library(skimr)            # Quick data summary
library(lubridate)        # Working with dates
library(readr)            # Reading CSV files
library(ggplot2)          # Visualization
library(ggpubr)           # Publication-ready plots
library(forcats)          # Categorical variables handling

# Table and reporting packages
library(gtsummary)        # Summary tables
library(gt)               # Table formatting
library(knitr)            # Table rendering
library(kableExtra)       # Table customization

# Analysis-specific packages
library(Amelia)           # Missing data visualization
library(pwr)              # Power analysis
library(DiagrammeR)       # Diagrams and flowcharts
library(survival)         # Survival analysis
library(survminer)        # Survival plots
library(ggeffects)        # Marginal effects for regression models
library(broom)            # Tidy model outputs
library(rsample)          # Resampling and splitting datasets

```

*Data Import*
```{r}
# Data Import & Initial Inspection 
# load the Dataset
PROMO_Data <- read_csv(here("data", "raw-data", "PROMO_Data.csv"))

promo_data_clean <- read.csv(here("data", "clean", "PROMO_Data_clean.csv"))
```

*Initial Inspection*

```{r}
# Check column names
colnames(promo_data_clean)

# Basic structure
str(promo_data_clean)

# Quick look at first few rows
head(promo_data_clean)

# Overview of missing data
skimr::skim(promo_data_clean)

# Count total missing values per column
promo_data_clean %>% 
  summarise(across(everything(), ~sum(is.na(.))))

```

```{r}
colnames(promo_data_clean)
```

```{r}
names(promo_data_clean)
```

*Summary Statistics*
Numeric Variables
```{r}
# Summary of numeric variables
promo_data_clean %>% 
  select(where(is.numeric)) %>% 
  summary()

```

*Categorical Variables*
```{r}
# Frequency tables for categorical variables
promo_data_clean %>% 
  select(where(is.character)) %>% 
  map(~table(.))

```

*Missing Data Visualization*
```{r}
# Visualizing missing data pattern
naniar::vis_miss(promo_data_clean)

```

*Univariate Plots*
Numeric Variables
```{r}
# Histograms of numeric variables
promo_data_clean %>% 
  select(where(is.numeric)) %>% 
  gather() %>% 
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "black") +
  facet_wrap(~key, scales = "free") +
  theme_minimal()

```

*Categorical Variables*
```{r}
# Bar plots of categorical variables
promo_data_clean %>% 
  select(where(is.character)) %>% 
  gather() %>% 
  ggplot(aes(x = value)) +
  geom_bar(fill = "lightblue", color = "black") +
  facet_wrap(~key, scales = "free") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

*Correlation Matrix (Numeric Variables)*
```{r}
# Correlation plot
numeric_vars <- promo_data_clean %>% 
  select(where(is.numeric))

cor_matrix <- cor(numeric_vars, use = "pairwise.complete.obs")

corrplot::corrplot(cor_matrix, method = "circle", type = "lower")

```

*Group-wise Summary*
```{r}
# Grouped summary by treatment group (replace 'treatment_group' with actual variable)
promo_data_clean %>% 
  group_by(treatment_group) %>% 
  summarise(across(where(is.numeric), list(mean = mean, sd = sd), na.rm = TRUE))

```